---
title: "Mini Project 1A: Static Maps"
output: html_document
date: "2024-09-20"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(mdsr)
library(maps)
library(dplyr)
library(plotly)
library(sf)


```


1. A Static Numeric Map: Total Violent Crime Rate by State(Excluding Alaska and Hawaii)

```{r}
# Load necessary libraries
library(ggplot2)
library(maps)
library(dplyr)


crime <- read.csv("https://corgis-edu.github.io/corgis/datasets/csv/state_crime/state_crime.csv")


crime_data <- crime %>%
  select(State, Data.Rates.Violent.All)


crime_data <- crime_data %>%
  filter(!State %in% c("Alaska", 
                       "Hawaii"))


crime_data$State <- tolower(crime_data$State)

# US map data
us_states <- map_data("state")

# Merging data
map_data_merged <- merge(us_states, 
                         crime_data, 
                         by.x = "region", 
                         by.y = "State", 
                         all.x = TRUE)

# Creating the map
ggplot(map_data_merged, 
       aes(x = long, 
           y = lat, 
           group = group, 
           fill = Data.Rates.Violent.All)) +
  geom_polygon(color = "black") + 
  scale_fill_continuous(name = "Violent Crime Rate", 
                        low = "lightyellow", 
                        high = "darkred", 
                        na.value = "grey") +
  theme_minimal() +
  labs(title = "Total Violent Crime Rate by State (Excluding Alaska & Hawaii)", 
       x = "Longitude", 
       y = "Latitude",
       caption = "Source: Corgis State Crime Dataset") +
  theme(plot.title = element_text(hjust = 0.5, 
                                  face = "bold", 
                                  size = 14),
        plot.subtitle = element_text(hjust = 0.5, 
                                     size = 12))

```


Alt-text:
This is a static map that visualizes the total violent crime rate by state in the United States, excluding Alaska and Hawaii. Each state is colored according to its violent crime rate per 100,000 people, with a gradient scale transitioning from light yellow (lower crime rates) to dark red (higher crime rates). The plot is titled "Total Violent Crime Rate by State (Excluding Alaska and Hawaii)," and there is a legend on the side indicating the color scale for crime rates. An annotation at the bottom notes the data source as the CORGIS State Crime Dataset. 

This map makes it easy to identify areas with high and low crime rates and the purpose of it is to see the regional differences in violent crime rates which can be used in discussions of public safety and law policies. States colored in darker shades of red indicate a higher rate of violent crime, leading to potentially looking into the underlying causes. Additionally, the exclusion of Alaska and Hawaii allows for a focused comparison among the contiguous states. We can see that all states are of varied shades of light yellow, which is not that helpful when it comes to figuring out the state with the highest and lowest. This could be improved by looking at violent crime rates in a specific time span.




2. A Static Categorical Map: Most Common Violent Crime by State (Excluding Alaska & Hawaii
ˆ
```{r}
# Load necessary libraries
library(ggplot2)
library(maps)
library(dplyr)

# Load the crime dataset
crime <- read.csv("https://corgis-edu.github.io/corgis/datasets/csv/state_crime/state_crime.csv")

# Exclude Alaska and Hawaii and select relevant columns
violent_crimes <- crime %>%
  filter(!State %in% c("Alaska", "Hawaii")) %>%
  select(State, Data.Totals.Violent.Assault, Data.Totals.Violent.Rape, 
         Data.Totals.Violent.Robbery, Data.Totals.Violent.Murder)


# Convert state names to lowercase for joining with map data
violent_crimes <- violent_crimes %>%
  mutate(State = tolower(State))

# Find the most common violent crime type for each state
violent_crimes <- violent_crimes %>%
  mutate(Highest_Crime_Type = case_when(
    Data.Totals.Violent.Assault >= Data.Totals.Violent.Rape & 
    Data.Totals.Violent.Assault >= Data.Totals.Violent.Robbery & 
    Data.Totals.Violent.Assault >= Data.Totals.Violent.Murder ~ "Assault",
    
    Data.Totals.Violent.Rape >= Data.Totals.Violent.Assault & 
    Data.Totals.Violent.Rape >= Data.Totals.Violent.Robbery & 
    Data.Totals.Violent.Rape >= Data.Totals.Violent.Murder ~ "Rape",
    
    Data.Totals.Violent.Robbery >= Data.Totals.Violent.Assault & 
    Data.Totals.Violent.Robbery >= Data.Totals.Violent.Rape & 
    Data.Totals.Violent.Robbery >= Data.Totals.Violent.Murder ~ "Robbery",
    
    Data.Totals.Violent.Murder >= Data.Totals.Violent.Assault & 
    Data.Totals.Violent.Murder >= Data.Totals.Violent.Rape & 
    Data.Totals.Violent.Murder >= Data.Totals.Violent.Robbery ~ "Murder",
    
    TRUE ~ NA_character_
  ))

# Load US map data
us_states <- map_data("state")

# Merge map data with crime data
map_data_merged <- merge(us_states, violent_crimes, by.x = "region", by.y = "State", all.x = TRUE)

# Create a categorical static map
ggplot(map_data_merged, aes(x = long, y = lat, group = group, fill = Highest_Crime_Type)) +
  geom_polygon(color = "black") + 
  scale_fill_manual(values = c("Assault" = "lightblue", "Rape" = "antiquewhite2", "Robbery" = "cornflowerblue", "Murder" = "burlywood3"), 
                    na.value = "grey", name = "Most Common Crime Type") +
  theme_minimal() +
  labs(title = "Most Common Violent Crime by State (Excluding Alaska & Hawaii)",
       subtitle = "Categorical Map of the U.S. Showing the Most Prevalent Type of Violent Crime",
       x = "Longitude", 
       y = "Latitude",
       caption = "Source: Corgis State Crime Dataset") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        legend.title = element_text(face = "bold"))

```

Description:
For this specific map, I intended it to be categorical but not interactive. This is an easy map to show the most common violent crime act in every state where red is assault, blue is rape and green is robbery. We can see that assault dominates most of the united states while robbery(green) dominates 7 states only. There is zero domination of rape crimes. I like this map since it's very straight forward, I would want to add abbreviations of each state to make it easier for the viewers to detect where every crime is dominant.

